Inference with PyroVelocity

### 2025-05-27 nbs/guides/inference.qmd
docs: preprocess data in inference guide
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

```python
import anndata as ad
import numpy as np
from pyrovelocity.models import PyroVelocity
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()

@@

# Create model
model = PyroVelocity(
    adata,
    model_type="normal",  # Options: "normal", "deterministic", "stochastic"
    guide_type="auto",    # Options: "auto", "delta"
    likelihood="Poisson", # Options: "Poisson", "NegativeBinomial"
    shared_time=True,     # Whether to use shared time across genes
    t_scale_on=False      # Whether to use time scaling
)

# Train model
model.train(
    max_epochs=1000,
    learning_rate=0.01,
    use_gpu="auto"
)

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=100
)

@@

# Save model
model.save_model("path/to/save/model")

# Load model
model = PyroVelocity.load_model("path/to/save/model", adata)
```

@@

```python
import anndata as ad
import numpy as np
from pyrovelocity.models.modular import (
    create_legacy_model2,
    PyroVelocityModel
)
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()

@@

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=100
)

# Store results in AnnData
adata = model.store_results_in_anndata(
    adata=adata,
    posterior_samples=posterior_samples
)
```

>>>>

```python
import anndata as ad
import numpy as np
from pyrovelocity.models import PyroVelocity
from pyrovelocity.tasks.preprocess import preprocess_dataset
from pyrovelocity.io.datasets import bifurcation_14

# Load data
adata = bifurcation_14()
adata, _, _ = preprocess_dataset("bifurcation_14", adata)

@@

# Create model
model = PyroVelocity(
    adata=adata,
    likelihood="Poisson",
    model_type="auto",
    guide_type="auto",
    correct_library_size=True,
    add_offset=True,
)

# Train model
model.train_faster(
    max_epochs=1000,
    lr=0.01,
    use_gpu="auto"
)

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=5,
)

@@

# Save model
model.save_model("models/bifurcation_14_model2/model")

# Load model
model = PyroVelocity.load_model("models/bifurcation_14_model2/model", adata)
```

@@

```python
import anndata as ad
import numpy as np
from pyrovelocity.models.modular import (
    PyroVelocityModel
)
from pyrovelocity.models.modular.factory import create_legacy_model2
from pyrovelocity.tasks.preprocess import preprocess_dataset
from pyrovelocity.io.datasets import bifurcation_14
from pyrovelocity.utils import print_anndata

# Load data
adata = bifurcation_14()
adata, _, _ = preprocess_dataset("bifurcation_14", adata)

@@

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=5,
)

# Store results in AnnData
adata = model.store_results_in_anndata(
    adata=adata,
    posterior_samples=posterior_samples
)
print_anndata(adata)
```
